---
title: "Introduction to mr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to mr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(mr)
```


## Quick Introduction

You create musical scores easily with mr.

Let's go through an example.

```{r}
m <- Music()
```

The above code initialized a `Music` object, to which we can add components:

```{r}
m <- m +
  Meter(4, 4) +
  Line(pitches = list("C5", "E5"), durations = list(2, 2))
```

We have added two components to the `Music` object:

1. a 4/4 time signature and
2. a musical line consisting of a C5 half note and an E5 half note.

If we print the `Music` object in an R console, we get a brief description of
it:

```{r}
m
```

We can convert the `Music` object to a musical score and an audio file:

```{r}
show(m, to = c("score", "audio"))
```

We can continue to add components and convert the `Music` object.

Add a tempo mark, for example:

```{r}
m <- m + Tempo(120)
show(m)
```

Add another musical line as a part:

```{r}
m <- m + Line(
  pitches = list("C3", "D3", "E3", "F3"),
  durations = list(1, 1, 1, 1)
)

show(m)
```

Add another musical line as a voice:

```{r}
m <- m + Line(
  pitches = list(c("E4", "G4")),
  durations = list(4),
  as = "voice",
  to = 1
)

show(m)
```

So now you have a rough idea of what mr does, before we go deeper, let's
learn how to install and configure it.


## Installation and Configuration

Install mr:

```r
install.packages("mr")
```

MuseScore, an open source and free notation software, is required. Get it
from <https://musescore.org/>.

If you don't install MuseScore to the default path (see
<https://musescore.org/en/handbook/3/revert-factory-settings>),
you need to tell mr where to find MuseScore:

Open ".Renviron" file, add `MUSESCORE_PATH=/your/path/to/musescore` into
it. You can open the file with `usethis::edit_r_environ()`.
